---
title: "ML_Homework1"
author: "John Navarro"
date: "June 23, 2017"
output: pdf_document
---
Data for the project are contained in two files: test_sample_X.csv and test_sample_Y.csv.
File test_sample_X.csv has 500 columns X1,X2,.,X500 of regressors. The corresponding column names are: X1, X2,., X500.

File test_sample_Y.csv contains 499 columns of dependent variables Y2,Y3,.,Y500Y2,Y3,.,Y500 with column names Y2, Y3,., Y500.

Each column Yj is the dependent variable of model mj with j regressors X1,X2,.,Xj
For example, second column Y3 is the dependent variable of the following model m3 with columns of predictors X1, X2, X3:
Y3=??0+??1X1+??2X2+??3X3+??.

These data are structured as in workshop 1 of this week.

Download your sample from left sidebar unpack and read it.
```{r}
dataPath <- "C:/Users/JohntheGreat/Documents/MSCA/MachineLearning/Week1_Regression_MLE"
test_data_regressors <- read.table(paste(dataPath,'test_sample_X.csv',sep = '/'), header=TRUE)
test_data_outputs <- read.table(paste(dataPath,'test_sample_Y.csv',sep = '/'), header=TRUE)

```

For each j=2,3,…,490 fit linear model mjmj.
For each mj, j=2,3,…,490 calculate 95% confidence interval for slope β1 of predictor  X1
Create two-column matrix intervals with 489 rows in which row j contains confidence interval of coefficient1β1 in model mj+1.

Create data frame res of two variables left and right

```{r}
left<-suppressWarnings(sapply(2:490,function(z) confint(lm(Y~.,data=data.frame(Y=test_data_outputs[,z-1],test_data_regressors[,1:z])))[2,1]))

right<-suppressWarnings(sapply(2:490,function(z) confint(lm(Y~.,data=data.frame(Y=test_data_outputs[,z-1],test_data_regressors[,1:z])))[2,2]))

# Create a matrix called intervals
intervals <- as.matrix(cbind(left, right))

# Create a dataframe "res", each row represents the left and right conf interval of each model
res<-data.frame(left=intervals[,1],right=intervals[,2])

# Save to a .csv file
write.csv(res,file=paste(dataPath,"W1answer.csv",sep="/"),row.names = F,quote=F)
```

