---
title: "Workshop"
author: "John Navarro"
date: "June 30, 2017"
output: pdf_document
---
```{r}

```

```{r}
set.seed(8394756)
Epsilon<-rnorm(500,0,1)
X<-rnorm(500*500,0,2)
dim(X)<-c(500,500)
colnames(X)<-paste0("X",1:500)
slopesSet<-runif(500,1,3)
Y<-sapply(2:500,function(z) 1+X[,1:z]%*%slopesSet[1:z]+Epsilon)
head(X[,1:5])
```

Fit LM
```{r}
m10<-lm(Y~.,data=data.frame(Y=Y[,9],X[,1:10]))
completeModelDataFrame<-data.frame(Y=Y[,490],X[,1:491])
m490<-lm(Y~.,data=completeModelDataFrame)
```

Relative measures
```{r}
install.packages("relaimpo")
library(relaimpo)
(metrics10 <- calc.relimp(m10, type= c("lmg","first","last","betasq", "pratt")))
```
Rank 10 predictors using method lmg
Reorder predictors according to method lmg

```{r}
(metrics10.lmg.rank<-metrics10@lmg.rank)
```
```{r}
orderedPredictors <- cbind(Predictors=colnames(X[,1:10])[order(metrics10.lmg.rank)],Slopes=slopesSet[1:10][order(metrics10.lmg.rank)])

  
originalR2.10<-sapply(2:10,function(z) summary(lm(Y~.,data=data.frame(Y=Y[,9],X[,1:z])))$r.squared)
improvedR2.10<-sapply(2:10,function(z) summary(lm(Y~.,data=data.frame(Y=Y[,9],orderedPredictors$Slopes[,1:z])))$r.squared)
matplot(2:10,cbind(originalR2.10,improvedR2.10),type="l",lty=1,lwd=2,col=c("black","red"),
        main="Improvement of Fit with Number of Predictors",
        xlab="Number of Predictors",ylab="Determination Coefficient")
legend("bottomright",legend=c("Original","Improved"),lty=1,lwd=2,col=c("black","red"))  
```

# 2. Method of PCA

Relplace original 10 predictors w their Principal components

```{r}
metrics10.lmg
PCA.10 <- princomp(data.frame(X[,1:10]))
summary(PCA.10)
#Do PCA on predictors only. Then reduce the # of predictors by replacing w most important factors. 
```
Extract factorLoadings, factorScores, zeroLoading
```{r}
factors10Data<-data.frame(Y=Y[,9],factorScores)
m10.PCA<-lm(Y~.,data=factors10Data)
#look at summary of m10.PCA and m10
```

# Fit lm to 491 predictors
make pCA for491 predictors w no response
look at importane graph of factors
definefactors/loadings/zero loadings
fit lm w PCA factors
get relative imporance
use prcomp()

```{r}
# build data frame, with 490th column as y, and 
completeModelDataFrame<-data.frame(Y=Y[,490],X[,1:491])
#run lm on dataset
m490<-lm(Y~.,data=completeModelDataFrame)
#run PCA for 491 predictors w no response
PCA.491 <- prcomp(X[,1:491])
#calculate importance graph
impo <- PCA.491$sdev/(sum(PCA.491$sdev))
plot(impo, type="l")
#extract factors and rotation
PCA.491.factors <- PCA.491$x
PCA.491.rotation <- PCA.491$rotation
factors491Data <- data.frame(Y=Y[,490], PCA.491.factors)
m491.PCA <- lm(Y~., factors491Data)
(metrics.m491.PCA <- calc.relimp(m491.PCA, type=c("first")))
sum.PCA.first=sum(metrics.m491.PCA@first)
sum.PCA.first
```
rank the factors
```{r}
(first.PCA.rank <- metrics.m491.PCA@first.rank)
orderedFactors<-factorScores[,order(first.PCA.rank)]
```




# Workshop 3 Labs from book
1. PCR

```{r}
suppressWarnings(library (pls))
suppressWarnings(library (ISLR))
data(Hitters)
head(Hitters)
```

