---
title: "ML Week 8 Workshop 3"
author: "John Navarro"
date: "August 11, 2017"
output: pdf_document
---

# Case Study: Spam

# 1 Spam Data

Spam data were collected by Hewlett-Packard Labs. The set contains 4601 emails of types "spam", "non-spam".
Besides the type in the file there are 57 predictors indicating frequencies of words and characters from vocabulary.
Help file explains the column meanings:

The first 48 variables contain the frequency of the variable name (e.g., business) in the e-mail. If the variable name starts with num (e.g., num650) the it indicates the frequency of the corresponding number (e.g., 650). The variables 49-54 indicate the frequency of the characters ';', '(', '[', '!', '$', and '#'. The variables 55-57 contain the average, longest and total run-length of capital letters.

Data are available through library kernlab.
```{r}
# install libraries
suppressWarnings(library(e1071))
suppressWarnings(library(kernlab))
```

```{r}
# read in spam data
data(spam)

# print the head of spam
head(spam,2)

# check the dimensions of spam
dim(spam) # 4601 X 58
```

Prepare train and test samples

```{r}
# assign number of observations to value N
N<-length(spam$type)

# assign nTrain to be 2/3s of the data set
nTrain<-round(2*N/3)

# set seed for reproducibility
set.seed(19)

# create an index to sample spam
idxShuffle<-sample(1:N,replace=F)

# subset spam by the index
spam_sample<-spam[idxShuffle,]

# separate the sample int training set
spam_train<-spam_sample[1:nTrain,]

# and testing set
spam_test<-spam_sample[(nTrain+1):N,]
```

# 2. Fitting SVM with default kernel

Tune SVM with radial kernel (default) like in workshop example with Galton's data.
Use grid for gamma= 10^(-6:-3) and cost= 10^(1:2)

```{r}
# set seed for reproducibility
set.seed(1)

# use svm to fit training data
svm.fit <- svm(type~., data=spam_train)

# Print summary of the svm model fit
summary(svm.fit)

# set seed for reproducibility
set.seed(1)

#grid search
svmTuned <- tune.svm(type~., data = spam_train, gamma = 10^(-6:-3), cost = 10^(1:2))

# summary of tuned svm model
summary(svmTuned)
```

```{r}

# plot the model's performance of gamma vs cost

#plot(m, transform.x=log10, xlab="expression",
      (log(10)(gamma)), ylab="C")

```


```{r}
# Print the best parameters given by the tuned model
svmTuned$best.parameters

# print the summary of the best model
summary(svmTuned$best.model)

# Predict classification on the spam test set using the best model
pred <- predict(svmTuned$best.model, spam_test)

# Print the confusion matrix
(acc <- table(pred, spam_test$type))

# print the accuracy
classAgreement(acc)

```


# 3 Experimenting with kernels

```{r}
# Fit svm with sigmoid kernel
svm.fit.sig <- svm(type~., data=spam_train, kernel="sigmoid" )
summary(svm.fit.sig)

#tune the sigmoid model
svmTuned.sig <- tune.svm(type~., data = spam_train, gamma = 10^(-6:-3), cost = 10^(1:2))
summary(svmTuned.sig)

# return the best model and create predictions, measure accruacy
svmTuned.sig$best.parameters
summary(svmTuned.sig$best.model)
pred <- predict(svmTuned.sig$best.model, spam_test)
(acc <- table(pred, spam_test$type))
```

```{r}
# linear kernael

svm.fit.linear <- svm(type~., data=spam_train, kernel="linear" )
summary(svm.fit.linear)

svmTuned.linear <- tune.svm(type~., data = spam_train, gamma = 10^(-6:-3), cost = 10^(1:2))
summary(svmTuned.linear)

svmTuned.linear$best.parameters
summary(svmTuned.linear$best.model)
pred <- predict(svmTuned.linear$best.model, spam_test)
(acc <- table(pred, spam_test$type))
classAgreement(acc)
```


```{r}
# polynomieal kernel deg 3

svm.fit.poly <- svm(type~., data=spam_train, kernel="polynomial", degree=3 )
summary(svm.fit.poly)

svmTuned.poly <- tune.svm(type~., data = spam_train, gamma = 10^(-6:-3), cost = 10^(1:2))
summary(svmTuned.poly)

svmTuned.poly$best.parameters
summary(svmTuned.poly$best.model)
pred <- predict(svmTuned.poly$best.model, spam_test)
(acc <- table(pred, spam_test$type))
```



```{r}
# polynomieal kernel deg 2

svm.fit.poly <- svm(type~., data=spam_train, kernel="polynomial", degree=4 )
summary(svm.fit.poly)

svmTuned.poly <- tune.svm(type~., data = spam_train, gamma = 10^(-6:-3), cost = 10^(1:2))
summary(svmTuned.poly)

svmTuned.poly$best.parameters
summary(svmTuned.poly$best.model)
pred <- predict(svmTuned.poly$best.model, spam_test)
(acc <- table(pred, spam_test$type))
```



```{r}

```

